---
title: "5. Geo-location Summary"
---

<hr>

```{r}
#| include: false
source("C:/Users/AYOMIDE/Documents/R/R_Projects/eda_SQL/function.R")
```

```{r}
#| message: false
#| warning: false
#| include: false
library(RMariaDB)
library(DBI)
library(gt)
library(gtExtras)
library(dplyr)
```


```{r}
#| message: false
#| warning: false
#| include: false
mysqldb <- RMariaDB::dbConnect(
  RMariaDB::MariaDB(), user =  "root", 
  host = "localhost",
  password = Sys.getenv("password"), 
  dbname = "sales"
  )
```

##### A temporary table containing sales, cost and profit. 
```{sql}
#| eval: false
CREATE TEMPORARY TABLE revenue_tbl
      SELECT store_id,
             year(order_date) AS order_year,
             (order_quantity * unit_price) AS sales,
             ((order_quantity * unit_cost) + ((order_quantity * unit_price) * discount_applied)) AS cost,
             (order_quantity * unit_price) - ((order_quantity * unit_cost) + ((order_quantity * unit_price) * discount_applied)) AS profit
        FROM sales_order;
```

```{r}
#| echo: false
#| warning: false
#| message: false
query <- "
CREATE TEMPORARY TABLE revenue_tbl
  	SELECT store_id,
           year(order_date) AS order_year,
           (order_quantity * unit_price) AS sales,
  		     ((order_quantity * unit_cost) + ((order_quantity * unit_price) * discount_applied)) AS cost,
  		     (order_quantity * unit_price) - ((order_quantity * unit_cost) + ((order_quantity * unit_price) * discount_applied)) AS profit
  		FROM sales_order;
"
results <- dbSendQuery(mysqldb, query)
dbClearResult(results)
```

### Total sales, cost and profit By :-

::: {.panel-tabset}

## Region
```{sql}
#| eval: false
SELECT r.region, 
       round(SUM(rt.sales), 2) AS Total_Sales,
       round(SUM(rt.cost), 2) AS Total_Cost,
       round(SUM(rt.profit), 2) AS Total_Profit
  FROM revenue_tbl rt
  JOIN store s
  	ON rt.store_id = s.store_id
  JOIN regions r
  	ON s.state_code = r.state_code
  GROUP BY r.region
  ORDER BY Total_Profit DESC;
```

```{r}
#| echo: false
fetch_res(
  "
  SELECT r.region, 
	       round(SUM(rt.sales), 2) AS Total_Sales,
         round(SUM(rt.cost), 2) AS Total_Cost,
         round(SUM(rt.profit), 2) AS Total_Profit
  	FROM revenue_tbl rt
    JOIN store s
  		ON rt.store_id = s.store_id
  	JOIN regions r
  		ON s.state_code = r.state_code
  	GROUP BY r.region
    ORDER BY Total_Profit DESC;
  "
) |>
  loc_sales_cost_profit_gt()
```

 
The highest profit was made in th west region, followed by the South and
Midwest, while the Northeast Region was where the lowest profit was made.


## State 
```{sql}
#| eval: false
SELECT s.state, 
       round(SUM(rt.sales), 2) AS Total_Sales,
       round(SUM(rt.cost), 2) AS Total_Cost,
       round(SUM(rt.profit), 2) AS Total_Profit
  FROM revenue_tbl rt
  JOIN store s
    ON rt.store_id = s.store_id
  GROUP BY s.state
  ORDER BY Total_Profit DESC;
```

```{r}
#| echo: false
fetch_res(
  "
  SELECT s.state, 
	       round(SUM(rt.sales), 2) AS Total_Sales,
         round(SUM(rt.cost), 2) AS Total_Cost,
         round(SUM(rt.profit), 2) AS Total_Profit
  	FROM revenue_tbl rt
       JOIN store s
  		   ON rt.store_id = s.store_id
  	GROUP BY s.state
    ORDER BY Total_Profit DESC;
  "
) |>
  loc_sales_cost_profit_gt()
```


## County
```{sql}
#| eval: false
SELECT s.county, 
       round(SUM(rt.sales), 2) AS Total_Sales,
       round(SUM(rt.cost), 2) AS Total_Cost,
       round(SUM(rt.profit), 2) AS Total_Profit
  FROM revenue_tbl rt
  JOIN store s
  	ON rt.store_id = s.store_id
  GROUP BY s.county
  ORDER BY Total_Profit DESC;
```

```{r}
#| echo: false
fetch_res(
  "
  SELECT s.county, 
	       round(SUM(rt.sales), 2) AS Total_Sales,
         round(SUM(rt.cost), 2) AS Total_Cost,
         round(SUM(rt.profit), 2) AS Total_Profit
  	FROM revenue_tbl rt
    JOIN store s
  		ON rt.store_id = s.store_id
  	GROUP BY s.county
    ORDER BY Total_Profit DESC;
  "
) |>
  loc_sales_cost_profit_gt()
```


## City
```{sql}
#| eval: false
SELECT s.city_name AS City, 
       round(SUM(rt.sales), 2) AS Total_Sales,
       round(SUM(rt.cost), 2) AS Total_Cost,
       round(SUM(rt.profit), 2) AS Total_Profit
  FROM revenue_tbl rt
  JOIN store s
  	ON rt.store_id = s.store_id
  GROUP BY s.city_name
  ORDER BY Total_Cost DESC;
```

```{r}
#| echo: false
fetch_res(
  "
  SELECT s.city_name AS City, 
	       round(SUM(rt.sales), 2) AS Total_Sales,
         round(SUM(rt.cost), 2) AS Total_Cost,
         round(SUM(rt.profit), 2) AS Total_Profit
  	FROM revenue_tbl rt
    JOIN store s
  		ON rt.store_id = s.store_id
  	GROUP BY s.city_name
    ORDER BY Total_Cost DESC;
  "
) |>
  loc_sales_cost_profit_gt()
```

:::


### Top 5 locations with the highest profit:

::: {.panel-tabset}

## overall
```{sql}
#| eval: false
SELECT r.region, s.state, s.county, s.city_name,
       round(SUM(rt.sales), 2) AS Total_Sales,
       round(SUM(rt.cost), 2) AS Total_Cost,
       round(SUM(rt.profit), 2) AS Total_Profit
  FROM revenue_tbl rt
  JOIN store s
  	ON rt.store_id = s.store_id
  JOIN regions r
  	ON s.state_code = r.state_code
  GROUP BY r.region, s.state, s.county, s.city_name
  ORDER BY Total_Profit DESC
  LIMIT 5;
```

```{r}
#| echo: false
fetch_res(
  "
  SELECT r.region, s.state, s.county, s.city_name,
	       round(SUM(rt.sales), 2) AS Total_Sales,
         round(SUM(rt.cost), 2) AS Total_Cost,
         round(SUM(rt.profit), 2) AS Total_Profit
  	FROM revenue_tbl rt
    JOIN store s
  		ON rt.store_id = s.store_id
  	JOIN regions r
  		ON s.state_code = r.state_code
  	GROUP BY r.region, s.state, s.county, s.city_name
    ORDER BY Total_Profit DESC
    LIMIT 5;
  "
) |>
  cl_name() |>
  gt() |>
  fmt_number(columns = 5:7) |>
  tbl_opts() |>
  cols_merge(columns = c(Region, State, County, `City Name`),
             pattern = "{1} Region, {2} State, {3}<br>{4}") |>
  cols_label(Region = "Location") |>
  data_color(columns = 5:6, colors = "#999999", apply_to = "text") 
```


## 2020
```{sql}
#| eval: false
SELECT r.region, s.state, s.county, s.city_name,
       round(SUM(rt.sales), 2) AS Total_Sales,
       round(SUM(rt.cost), 2) AS Total_Cost,
       round(SUM(rt.profit), 2) AS Total_Profit
  FROM revenue_tbl rt
  JOIN store s
  	ON rt.store_id = s.store_id
  JOIN regions r
  	ON s.state_code = r.state_code
  WHERE order_year = 2020
  GROUP BY r.region, s.state, s.county, s.city_name
  ORDER BY Total_Profit DESC
  LIMIT 5;
```

```{r}
#| echo: false
fetch_res(
  "
  SELECT r.region, s.state, s.county, s.city_name,
	       round(SUM(rt.sales), 2) AS Total_Sales,
         round(SUM(rt.cost), 2) AS Total_Cost,
         round(SUM(rt.profit), 2) AS Total_Profit
  	FROM revenue_tbl rt
    JOIN store s
  		ON rt.store_id = s.store_id
  	JOIN regions r
  		ON s.state_code = r.state_code
    WHERE order_year = 2020
  	GROUP BY r.region, s.state, s.county, s.city_name
    ORDER BY Total_Profit DESC
    LIMIT 5;
  "
) |>
  cl_name() |>
  gt() |>
  fmt_number(columns = 5:7) |>
  tbl_opts() |>                       
  cols_merge(columns = c(Region, State, County, `City Name`),
             pattern = "{1} Region, {2} State, {3}<br>{4}") |>
  cols_label(Region = "Location") |>
  data_color(columns = 5:6, colors = "#999999", apply_to = "text") 
```

:::

<h3 class="insight-text">Insight</h3>

<p class="insight-summary">
In 2020 The store at Vallejo in Solano County in the state of California within the West region made the most profit for the company. While for all the past years including 2020 Broken Arrow store at Wagoner County in Oklahoma within the Southern region of the country made the most profit.
<p/>