---
title: "4. Method Of Product Sales"
---

<hr>


```{r}
#| include: false
source("C:/Users/AYOMIDE/Documents/R/R_Projects/eda_SQL/function.R")
```

```{r}
#| message: false
#| warning: false
#| include: false
library(RMariaDB)
library(DBI)
library(gt)
library(gtExtras)
library(dplyr)
```


```{r}
#| message: false
#| warning: false
#| include: false
mysqldb <- RMariaDB::dbConnect(
  RMariaDB::MariaDB(), user =  "root", 
  host = "localhost",
  password = Sys.getenv("password"), 
  dbname = "sales"
  )
```


### Starting with the unique sales channels and the number of transactions done through them.
```{sql}
#| eval: false
SELECT sales_channel, count(sales_channel) AS Count
  FROM sales_order
  GROUP BY sales_channel
  ORDER BY Count DESC;
```

```{r}
#| echo: false
fetch_res(
  "
  SELECT sales_channel, count(sales_channel) AS Count
	  FROM sales_order
    GROUP BY sales_channel
    ORDER BY Count DESC;
  "
) |>
  mutate(Count = as.integer(Count)) |>
  cl_name() |>
  gt() |>
  fmt_number(columns = 2, decimals = 0) |>
  tbl_opts()
```

 
Most sales can in through the physical store, followed by online platforms
and distributors, while the wholesale channel brought in the lowest number of
transactions.


### The total & average order quantity and sale from each sales channel.
```{sql}
#| eval: false
WITH sc_qos AS (
    SELECT sales_channel,
           order_quantity, 
           (order_quantity * unit_price) AS sales
      FROM sales_order
)
SELECT sales_channel,
       round(AVG(order_quantity), 2) AS Average_Order,
       SUM(order_quantity) AS Total_order,
       round(AVG(sales), 2) AS Average_Sales,
       round(SUM(sales), 2) AS Total_Sales
  FROM sc_qos
  GROUP BY sales_channel
  ORDER BY Total_Sales DESC;
```

```{r}
#| echo: false
f <- fetch_res(
  "
  WITH sc_qos AS (
    	SELECT sales_channel,
    		     order_quantity,
    		     (order_quantity * unit_price) AS sales
    		FROM sales_order
)
SELECT sales_channel,
	     round(AVG(order_quantity), 2) AS Average_Order,
       SUM(order_quantity) AS Total_order,
       round(AVG(sales), 2) AS Average_Sales,
       round(SUM(sales), 2) AS Total_Sales
	FROM sc_qos
  GROUP BY sales_channel
  ORDER BY Total_Sales DESC;
  "
)

f |>
  cl_name() |>
  gt() |>
  fmt_number(columns = c(2, 4)) |>
  fmt_number(columns = c(3, 5), decimals = 0) |>
  tbl_opts() |>
  gt_highlight_cols(columns = c(2, 4), alpha = 0.08, fill = "#F5F5F5") |>
  gt_highlight_cols(columns = c(3, 5), alpha = 0.3, fill = "#F2F2F2") |>
  gt_highlight_rows(columns = 2, rows = 4, fill = "#FAFAFA") |>
  gt_highlight_rows(columns = c(3, 5), rows = 1, fill = "#F2F2F2") |>
  gt_highlight_rows(columns = 4, rows = 3, fill = "#FAFAFA")
```


<h3 class="insight-text">Insight</h3>

<p class="insight-summary"> 
Given the rank of the number of transactions from each sales channel, the same
can also be said for the total amount of order and sales, with the exception
of the average sales of which the distributor channel had the highest while
the online channel had the lowest.
</p>